---
title: "STEM Completion Study"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load libraries

```{r}
library(tidyverse)
library(magrittr)
library(janitor)
library(kutils) #enables use of mgsub
library(dlookr)
library(flextable) 
library(vegan) # diversity measures

```
Read data into base dataframe (data) 

```{r}
data <- read.csv("C:/Users/seanc/Documents/Data_Projects/STEM_Completion_Study/Data/IPED_Bach_Res/10_11.csv")%>%as.data.frame()
```

create working df

```{r}
df<-data
```

Initial cleaning and review of the data

Results:

1. No duplicates found
2. Dim 2045, 272 - no empty rows or cells
3. There are 241 columns of integer type and 31 columns of character type
4. There are 30 columns allocated for imputation flags. These can be stored in a separate dataframe. 
```{r}
# clean names for consistency and lower case

df%<>%clean_names

# remove any empty rows and cols

df%<>%remove_empty(c("rows", "cols"))

# assess presence of duplicates

get_dupes(df)

# assess initial variable types and data dimensions

dim(df)
colnames(df)
str(df)

# sum column types with function
ncol_type <- function(df, ty){
  sum(sapply(df, function(x) class(x)==ty))
}

ncol_type(df, "integer")
ncol_type(df, "character")

# assess number of columns allocated imputation flags

length(grep(pattern = "imp", x = colnames(df)))  #grep searches for patterns
 

```

Create working dataframe (df) and rename columns to shorten

```{r}
#change col names for easier reading 

pattern<-c('unit_id', 'institution_name', 'grand_total','imputation_status_flag_for_', 'nonresident_alien', 'race_ethnicity_unknown', 'two_or_more_races', 'native_hawaiian_or_other_pacific_islander', 'hispanic_or_latino', 'black_or_african_american',  'american_indian_or_alaska_native', 'c2011_a_rv_first_major_', '_bachelor_s_degree', 'engineering', 'eng_technologies_and_eng_related_fields', 'computer_and_information_sciences_and_support_services','mathematics_and_statistics', 'physical_sciences','science_technologies_technicians','biological_and_biomedical_sciences', 'gran_tot_gran_tot', 'gran_tot_men_gran_tot', 'gran_tot_women_gran_tot')
 
abbrev<-c('id', 'name', 'gran_tot','imp_', 'nonres', 'unk', 'two', 'pacific', 'latino', 'black',  'native', '', '', 'eng', 'eng_tech', 'comp', 'math', 'phys', 'sci_tech', 'bio', 'gran_tot', 'gran_tot_men', 'gran_tot_women') 

name<-c(colnames(df))

subs <- mgsub(pattern, abbrev, name)  #note accomplished in order

colnames(df)<-subs
```

Assess Missing Data

1.98% of the schools are missing data for tech 
2.75% of the schools are missing data for engineering 
3.44% of the schools are missing data for physics
4.40% of the schools are missing data for math
5.38% of the schools are missing data for comp
6. 35% of the shools are missing data for biol
7. 9% of the schools are missing data for STEM completion grand totals

We will remove 60 cols for tech completion data 
We will remove 220 schools for that are missing STEM completion grand totals

Our dataframe dimensions are now 1825/212
```{r}
# assess missing by column (total and percentage)

df%>%
  diagnose()%>%
  select(-unique_count, -unique_rate)%>%
  filter(missing_count>0)%>%
  arrange(desc(missing_count))%>%
  flextable()

```
Remove data will high missingness as well as 0|NA grand totals for STEM completion


```{r}
# remove columns for tech programs due to high level of missingness

df%<>%select(!ends_with('tech'))

# remove schools that are missing STEM completion grand totals

df%<>%
    filter(gran_tot != 0 | gran_tot != NA )
```

Subset working dataframe for further manipulation and qaqc.

This will leave us with a data frame that tracks imputation flags(imputed_df)  and another 
that contains variable values (data_df)

```{r}
# subset imputed columns for later use if necessary

imputed_df<-df%>%
    dplyr::select(starts_with('imp'))

imputed_df<-cbind(imputed_df, df$id)%>%
  rename(id = 'df$id')

# subset data columns for feature engineering, EDA, and analyses - remove imputation, unk, nonres columns

data_df<-df%>%
    select(!starts_with(c('imp')))

```

Calc percentages for total by ethnicity and update col names

```{r}
ethnic<-data_df[3:32]

ethnic%<>%mutate(perc_men = (ethnic[[2]]/ethnic[[1]])*100, .after=gran_tot)%>%select(!(3))
ethnic%<>%mutate(perc_women = (ethnic[[3]]/ethnic[[1]])*100, .after=(2))%>%select(!(4))

ethnic%<>%mutate(perc_native = (ethnic[[4]]/ethnic[[1]])*100, .after=(3))%>%select(!(5))
ethnic%<>%mutate(perc_native_men = (ethnic[[5]]/ethnic[[1]])*100, .after=(4))%>%select(!(6))
ethnic%<>%mutate(perc_native_women = (ethnic[[6]]/ethnic[[1]])*100, .after=(5))%>%select(!(7))

ethnic%<>%mutate(perc_asian = (ethnic[[7]]/ethnic[[1]])*100, .after=(6))%>%select(!(8))
ethnic%<>%mutate(perc_asian_men = (ethnic[[8]]/ethnic[[1]])*100, .after=(7))%>%select(!(9))
ethnic%<>%mutate(perc_asian_women = (ethnic[[9]]/ethnic[[1]])*100, .after=(8))%>%select(!(10))

ethnic%<>%mutate(perc_black = (ethnic[[10]]/ethnic[[1]])*100, .after=(9))%>%select(!(11))
ethnic%<>%mutate(perc_black_men = (ethnic[[11]]/ethnic[[1]])*100, .after=(10))%>%select(!(12))
ethnic%<>%mutate(perc_black_women = (ethnic[[12]]/ethnic[[1]])*100, .after=(11))%>%select(!(13))

ethnic%<>%mutate(perc_latino = (ethnic[[13]]/ethnic[[1]])*100, .after=(12))%>%select(!(14))
ethnic%<>%mutate(perc_latino_men = (ethnic[[14]]/ethnic[[1]])*100, .after=(13))%>%select(!(15))
ethnic%<>%mutate(perc_latino_women = (ethnic[[15]]/ethnic[[1]])*100, .after=(14))%>%select(!(16))

ethnic%<>%mutate(perc_pacific = (ethnic[[16]]/ethnic[[1]])*100, .after=(15))%>%select(!(17))
ethnic%<>%mutate(perc_pacific_men = (ethnic[[17]]/ethnic[[1]])*100, .after=(16))%>%select(!(18))
ethnic%<>%mutate(perc_pacific_women = (ethnic[[18]]/ethnic[[1]])*100, .after=(17))%>%select(!(19))

ethnic%<>%mutate(perc_white = (ethnic[[19]]/ethnic[[1]])*100, .after=(18))%>%select(!(20))
ethnic%<>%mutate(perc_white_men = (ethnic[[20]]/ethnic[[1]])*100, .after=(19))%>%select(!(21))
ethnic%<>%mutate(perc_white_women = (ethnic[[21]]/ethnic[[1]])*100, .after=(20))%>%select(!(22))

ethnic%<>%mutate(perc_two = (ethnic[[22]]/ethnic[[1]])*100, .after=(21))%>%select(!(23))
ethnic%<>%mutate(perc_two_men = (ethnic[[23]]/ethnic[[1]])*100, .after=(22))%>%select(!(24))
ethnic%<>%mutate(perc_two_women = (ethnic[[24]]/ethnic[[1]])*100, .after=(23))%>%select(!(25))

ethnic%<>%mutate(perc_unk = (ethnic[[25]]/ethnic[[1]])*100, .after=(24))%>%select(!(26))
ethnic%<>%mutate(perc_unk_men = (ethnic[[26]]/ethnic[[1]])*100, .after=(25))%>%select(!(27))
ethnic%<>%mutate(perc_unk_women = (ethnic[[27]]/ethnic[[1]])*100, .after=(26))%>%select(!(28))

ethnic%<>%mutate(perc_nonres = (ethnic[[28]]/ethnic[[1]])*100, .after=(27))%>%select(!(29))
ethnic%<>%mutate(perc_nonres_men = (ethnic[[29]]/ethnic[[1]])*100, .after=(28))%>%select(!(30))
ethnic%<>%mutate(perc_nonres_women = (ethnic[[30]]/ethnic[[1]])*100, .after=(29))%>%select(!(31))


ethnic

```

Calculate percentages and change col names for major fields of study

```{r}
#create a list of dataframes based on field of study

fields_df<-c()

maj<-c("bio", "eng", "comp", "phys", "math")

for(i in seq_along(maj)){
    
    field<-data_df%>%select(c(gran_tot, ends_with(maj[[i]])))
    fields_df[[i]]<-field
    
}
# create function to calculate percentages for vars and rename col names

transform<-function(field, maj){
    
    new<-field
    new%<>%mutate(!!paste('perc_tot_', maj) := (field[[2]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_men_', maj) := (field[[3]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_women_', maj) := (field[[4]]/field[[1]])*100)
    
    new%<>%mutate(!!paste('perc_native_', maj) := (field[[5]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_native_men_', maj) := (field[[6]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_native_women_', maj) := (field[[7]]/field[[1]])*100)
    
    new%<>%mutate(!!paste('perc_asian_', maj) := (field[[8]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_asian_men_', maj) := (field[[9]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_asian_women_', maj) := (field[[10]]/field[[1]])*100)   
    
    new%<>%mutate(!!paste('perc_black_', maj) := (field[[11]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_black_men_', maj) := (field[[12]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_black_women_', maj) := (field[[13]]/field[[1]])*100)
    
    new%<>%mutate(!!paste('perc_latino_', maj) := (field[[14]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_latino_men_', maj) := (field[[15]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_latino_women_', maj) := (field[[16]]/field[[1]])*100)
    
    new%<>%mutate(!!paste('perc_pacific_', maj) := (field[[17]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_pacific_men_', maj) := (field[[18]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_pacific_women_', maj) := (field[[19]]/field[[1]])*100)
    
    new%<>%mutate(!!paste('perc_white_', maj) := (field[[20]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_white_men_', maj) := (field[[21]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_white_women_', maj) := (field[[22]]/field[[1]])*100)
    
    new%<>%mutate(!!paste('perc_two_total_', maj) := (field[[23]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_two_men_', maj) := (field[[24]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_two_women_', maj) := (field[[25]]/field[[1]])*100)
    
    new%<>%mutate(!!paste('perc_unk_total_', maj) := (field[[26]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_unk_men_', maj) := (field[[27]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_unk_women_', maj) := (field[[28]]/field[[1]])*100)
    
    new%<>%mutate(!!paste('perc_nonres_total_', maj) := (field[[29]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_nonres_men_', maj) := (field[[30]]/field[[1]])*100)
    new%<>%mutate(!!paste('perc_nonres_women_', maj) := (field[[31]]/field[[1]])*100)
    
    new%<>%mutate_all(~replace(., is.nan(.), 0))%>%
        select(c(32:61))
    
    print(new)
   
    return(new)
    
}

# execute function to update field dfs

df1<-transform(data.frame(fields_df[1]), 'bio')
df2<-transform(data.frame(fields_df[2]), 'eng')
df3<-transform(data.frame(fields_df[3]), 'comp')
df4<-transform(data.frame(fields_df[4]), 'phys')
df5<-transform(data.frame(fields_df[5]), 'math')

#create combined df

identifier<-data_df%>%select(1:2)

percent_df<-cbind(identifier, ethnic, df1,df2,df3,df4,df5)

```

# calculate shannon and simpson diversity indices


Simpson - The measure equals the probability that two entities taken at random from the dataset of interest represent the same type

```{r}

#subset dataframe to collect ethnic totals

totals_df<-data_df%>%select(c(ends_with('total_gran_tot')))%>%
    select(!c(unk_total_gran_tot, nonres_total_gran_tot))

shannon_df<-data.frame(diversity(totals_df, index = "shannon", MARGIN = 1, base = exp(1)))%>%
    rename(shannon_h = 'diversity.totals_df..index....shannon...MARGIN...1..base...exp.1..')%>%
    mutate(shannon_e = shannon_h/log(ncol(totals_df)))%>%
    mutate_if(is.numeric, round, digits = 2)

simpson_df<-data.frame(diversity(totals_df, index = "simpson", MARGIN = 1, base = exp(1)))%>%
    rename(simpson = 'diversity.totals_df..index....simpson...MARGIN...1..base...exp.1..')%>%
    mutate_if(is.numeric, round, digits = 2)

inv_simpson_df<-data.frame(diversity(totals_df, index = "inv", MARGIN = 1, base = exp(1)))%>%
    rename(inv_simpson = 'diversity.totals_df..index....inv...MARGIN...1..base...exp.1..')%>%
    mutate_if(is.numeric, round, digits = 2)

stemdiv_df<-cbind(identifier,shannon_df, simpson_df, inv_simpson_df)



```


```{r}

```



```{r}

```


```{r}

```



```{r}

```


```{r}

```



```{r}

```


```{r}

```

