---
title: 'STEM Gender Study - Data Cleaning'
date: "03/24/22"
author: "Sean Connin"
output:
  html_document: 
    toc: TRUE
    toc-title: ""
    toc_depth: 3
    toc_float: 
      collapsed: TRUE
      smooth_scroll: FALSE
    number_sections: true
    df_print: paged
    code_folding: hide
    theme: sandstone
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this project is to estimate the average percentage of women completing undergraduate degrees in STEM programs for public and private 4-YR+ institutions in the United States based on extrinsic factors. The latter include such covariates as institution size, location, sector, student:faculty ratio, etc. 

Data Source: Integrated Postsecondary Education Database System

The full dataset includes variables for 1803 post-secondary institutions (Public 4 Year +, Private non_profit) and comprises the period from 2010-2020, excluding 2011-12 and 2016-17. The latter are currently omitted due to data quality concerns. 

Included are the following disciplines: Computer Science, Biological Science, Engineering, Mathematics, and Physcial Science. 

This rmd files includes code for initial data cleaning and wrangling.

```{r}
library(tidyverse)
library(magrittr)
```

Load datasets, set initial col names, combine inst characteristics files

```{r}

gender<-read_csv("C:\\Users\\seanc\\OneDrive\\Documents\\Data_Projects\\STEM_Completion_Study\\Data\\IPED_Bach_Res\\Output\\totals.csv") # this dataset has issues with nonunique rows need to clean

gender%>%
    arrange(id, year)

sf.ratio.1920<-read_csv('C:\\Users\\seanc\\OneDrive\\Documents\\Data_Projects\\STEM_Completion_Study\\Data\\IPED_Bach_Res\\raw_data\\predictors_1920\\student_faculty_ratio_1920.csv')%>%
    select(c(UnitID, 'Student-to-faculty ratio (EF2019D_RV)'))%>%
    rename(sf_ratio = 'Student-to-faculty ratio (EF2019D_RV)')

ret.rate.1920<-read_csv("C:\\Users\\seanc\\OneDrive\\Documents\\Data_Projects\\STEM_Completion_Study\\Data\\IPED_Bach_Res\\raw_data\\predictors_1920\\full_time_retention_rate_1920.csv")%>%
    select(c(UnitID, 3))

colnames(ret.rate.1920)[2] <- 'ret_rate'  # could not get rename to work properly
  
perc.admit1920<-read_csv("C:\\Users\\seanc\\OneDrive\\Documents\\Data_Projects\\STEM_Completion_Study\\Data\\IPED_Bach_Res\\raw_data\\predictors_1920\\percent_admitted_1920.csv")%>%
    select(c(UnitID, 3:5))

colnames(perc.admit1920)[2] <- 'perc_wm_admit'  # could not get rename to work properly
colnames(perc.admit1920)[3] <- 'perc_mn_admit'  # could not get rename to work properly
colnames(perc.admit1920)[4] <- 'perc_admit'  # could not get rename to work properly

perc.enroll.1920<-read_csv("C:\\Users\\seanc\\OneDrive\\Documents\\Data_Projects\\STEM_Completion_Study\\Data\\IPED_Bach_Res\\raw_data\\predictors_1920\\percent_undergrad_enrollment_1920.csv")%>%
    select(c(UnitID, 12))
           
colnames(perc.enroll.1920)[2] <- 'perc_wm_enroll'  # could not get rename to work properly

inst1920<-left_join(sf.ratio.1920, ret.rate.1920, by = 'UnitID')
inst1920<-left_join(inst1920, perc.admit1920, by = 'UnitID')
inst1920<-left_join(inst1920, perc.enroll.1920, by = 'UnitID')

inst1920%<>%
    rename(id = 'UnitID')%>%
    relocate(c(ret_rate, sf_ratio), .after = perc_wm_enroll)%>%
    relocate(perc_admit, .after = id)%>%
    arrange(id)
    
```
subset data for analysis and join with inst1920 vars

```{r}

c.gender<-gender%>%
    select(c(1:14, ends_with('comp')))%>%
    select(!c(latitude, longitude, city))%>%
    mutate(perc_wm_comp = (gran_tot_women_comp/gran_tot_comp)*100)%>%
    mutate(perc_mn_comp = (gran_tot_men_comp/gran_tot_comp)*100)%>%
    mutate(perc_wm_stem = (tot_women/gran_tot)*100)%>%
    mutate(perc_mn_stem=(tot_men/gran_tot)*100)%>%
    select(c(1:11, 36:39))%>%
    relocate(c(perc_wm_stem, perc_mn_stem), .after = tot_women )%>%
    arrange(id, year)%>%
    distinct(.)%>%
    drop_na(4:6) 

stm.gender<-c.gender%>%    
    group_by(id)%>%
    select(c(9:15))%>%
    summarise(across(where(is.numeric), mean, na.rm= TRUE))%>%
    summarise(across(where(is.numeric),round))%>%
    arrange(id)
    
inst<-c.gender%>%
    select(1:8)%>%
    select(!year)%>%
    distinct()%>%
    arrange(id)
    
data.stm<-inner_join(inst, stm.gender, by = 'id')%>%arrange(id)

data.stm<-left_join(data.stm, inst1920)

# convert NaN to NA, 

d<-map(data.stm, ~ifelse(is.nan(.), NA, .))
data.stm<-data.frame(d)

# change char to factor

data.stm%<>%
    mutate_at(2:5, as.factor)%>%
    na.omit()

#write cleaned data to file 

write_csv(data.stm, 'gender_tree_data.csv')
```
